---
title: "p8105_hw3_zh2596"
author: "Zilin Huang"
date: "`r Sys.Date()`"
output: github_document
---

```{r, warning=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(readr)
library(dplyr)
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")
```

## Problem 1

This dataset shows the order status of a specific group of customers. with the orders' information and date recorded.

There are `nrow(instacart)` rows and `ncol(instacart)` columns in the `instacart` dataset. Some key variables include `user_id` that identifies unique users, `order_id` that identifies unique orders, `aisle_id`, `department_id` that identify the category that the product belongs to, etc.

As an example, on row 9, a customer with user id `79431` orders `Grated Pecorino Romano Cheese`, one of his/her eight orders on Saturday at 6 p.m. He/she has never ordered this product before.

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_obs = n()) |>
  arrange(desc(n_obs))
```

There are `r length(unique(instacart$aisle))` aisles, and based on the code chunk above, `fresh vegetables` contains the most items ordered from. 

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
This step limits the plot to aisles with more than 10000 items ordered.

```{r}
instacart |>
  group_by(aisle) |>
  mutate(n_obs = n()) |>
  arrange(desc(n_obs)) |>
  filter(n_obs > 10000) |>
  ggplot(aes(x = n_obs, fill = aisle)) + 
    geom_histogram()
```


### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>   
  filter(aisle %in% 
         c("baking ingrediens", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarize(n_obs=n())
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour=mean(order_hour_of_day)) |>
  pivot_wider(names_from=order_dow, values_from=mean_hour)
```

## Problem 2

Do some data cleaning for the `BRFSS` data:

```{r}
brfss_smart2010 |>
  filter(
    Topic == "Overall Health",
    Response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) |>
  mutate(
    Response == forcats::fct_relevel(Response, c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```

### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010 |>
  filter(Year == 2002) |>
  group_by(Locationabbr) |>
  summarise(number_of_loctions = n_distinct(Locationdesc)) |>
  filter(number_of_loctions >= 7)
```

```{r}
brfss_smart2010 |>
  filter(Year == 2010) |>
  group_by(Locationabbr) |>
  summarise(number_of_loctions = n_distinct(Locationdesc)) |>
  filter(number_of_loctions >= 7)
```

By the two tables above, the states abbreviated as "CT", "FL", "MA", "NC", "NJ", "PA" were observed at 7 or more locations inn 2002; while the states abbreviated as "CA", "CO", "FL", "MA", "MD", "NC", "NE", "NJ", "NY", "OH" were observed at 7 or more locations inn 2010.

### Construct a dataset that is limited to `Excellent` responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_smart2010_new =
  brfss_smart2010 |>
  filter(Response == "Excellent") |>
  select(Year, Locationabbr, Data_value) |>
  group_by(Year,Locationabbr) |>
  mutate(mean_data_value = mean(Data_value, na.rm=TRUE)) |>
  select(-Data_value) 

  ggplot(brfss_smart2010_new, aes(Year, mean_data_value, color=Locationabbr)) +
    geom_line() +
    labs(
      title="Average BRFSS Data Value Over Time among States",
      y="Average BRFSS Data Value"
    )
```
From the spaghetti plot, we can see that the average BRFSS data centered around the range of (15, 30) for all states, with some states containing minimal values (below 15) in different years.

### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
brfss_smart2010_filter =
  brfss_smart2010 |>
  filter(
    Year %in% c(2006, 2010),
    Response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor"),
    Locationabbr == "NY")

brfss_smart2010_filter |>
    ggplot(aes(x=Response,y=Data_value)) +
    facet_grid(~Year) + 
    geom_boxplot() +
    labs(
      title="Distribution of BRFSS Data Values in NY State",
      x="Response (from Poor to Excellent)",
      y="BRFSS Data Value"
    )
```

From the two-panel plot, we can see that the BRFSS data of different responses across the two years do not vary by a lot. On average, the "Good" response contains the highest data value, while the "Poor" response contains the lowest data value. 

## Problem 3


```{r}
# this step exclude participants less than 21 years of age, and those with missing demographic data
nhanes_covar =
  read_csv("./data/nhanes_covar.csv", skip=4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na() |>
  # this step encode data with reasonable variable classes
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"),
    sex = as.factor(sex),
    education =
      case_match(
        education,
        1 ~ "Less than high school",
        2 ~ "High school equivalent",
        3 ~ "More than high school"),
    education = as.factor(education),
    seqn = as.character(seqn)) 

nhanes_accel =
   read_csv("./data/nhanes_accel.csv") |>
   janitor::clean_names() |>
   mutate(seqn = as.character(seqn))
```


# this steps merges the two tables together
```{r}
merge_df =
  left_join(nhanes_covar, nhanes_accel, by = "seqn") |>
  pivot_longer(min1:min1440,
               values_to = "accelerometer data",
               names_to = "min")
```

### Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
# this step produces the table
nhanes_covar |>
  group_by(sex, education) |>
  summarize(number_of_male_or_female = n())
```

```{r}
# this step creates the plot of the age distributions
merge_df |>
  ggplot(aes(x=education,y=age,fill=sex)) +
  geom_boxplot() +
  labs(
    title="Age Distribution Acorss Sex and Education"
  )
```

This plot shows that the average age is larger for female receiving an education level of high school equivalent, while those who receive an education level of more than high school tend to be younger.

```{r}
merge_df =
  merge_df |>
  group_by(seqn) |>
  mutate(total_activity = sum(`accelerometer data`))

ggplot(data=merge_df, aes(x=age,y=total_activity, color=sex)) +
  geom_point() +
  geom_smooth() +
  facet_grid(~education) +
  labs(title="Accelerometer Data Trend across Education and Sex Groups")
```
The figure shows that in general, men have a lower value of accelerometer data than women when they receive high school education at least, while men's data values is higher than women in the group of "Less than High School". For all three groups of education, the data value has a decreasing trend when both men and women's age increase.

```{r}
ggplot(data=merge_df, aes(x=total_activity, fill=sex)) +
  geom_density(alpha=.5) +
  facet_grid(~education) + 
  viridis::scale_fill_viridis(discrete = TRUE) +
  labs(title="24-hour Activity Time Courses of Data across Education and Sex Groups")
```

The three-panel plot shows that the 24-hour activity time courses of the accelerometer data centers around the range of (10000, 15000) for each education and sex group. Specifically, female's data values has a more centralized focus than male in the group of "More than High School".